---
title: "Representation & Gender Diversity in Google’s Global Workforce"
output: html_notebook
---

## Guiding Research Question
- How does gender representation, hiring, and attrition vary across regions in Google's workforce?

## Data Access

**Primary Data**
- Region level workforce metrics such as representation, hiring, exits, and attrition.
- Source: Google Diversity Report from BigQuery & data available on this Github repo.
- - [Project Repository](https://github.com/PixelQuinn/Diversity-and-Inclusion-Analysis)

**Secondary Dataset:**
- Self-ID Representation data, including responses related to LGBTQ+ identity and disability status.
- Source: Internal Google workforce data (self-reported), accessed through Big Query.
- [Project Repository](https://github.com/PixelQuinn/Diversity-and-Inclusion-Analysis).

```{r}
library(ggplot2)
library(tidyverse)
library(readr)
library(tidycensus)
library(janitor)
library(dplyr)
library(ggrepel)
```

## Initial Inspection

```{r}
# Load data
region_rep <- read_csv("data-sets/region-non-intersectional-representation.csv") %>% clean_names()
region_hire <- read_csv("data-sets/region-non-intersectional-hiring.csv") %>% clean_names()
region_exit <- read_csv("data-sets/region-non-intersectional-exits-representation.csv") %>% clean_names()
region_attr <- read_csv("data-sets/region-non-intersectional-attrition-index.csv") %>% clean_names()

glimpse(region_rep)
summary(region_rep)
```

```{r}
self_ID <- read_csv("data-sets/selfid-representation.csv")

glimpse(self_ID)
```

```{r}
# Join the representation and hiring by region, year, and workforce type.

gender_rep_hiring <- region_rep %>%
  inner_join(region_hire, by = c("region", "report_year", "workforce"), suffix = c("_rep", "_hire"))
```

```{r}
gender_rep_hiring %>%
  filter(!is.na(gender_women_rep), !is.na(gender_women_hire)) %>%
  group_by(region) %>%
  summarise(
    avg_rep_women = mean(gender_women_rep),
    avg_hire_women = mean(gender_women_hire)
  )
```

```{r}
# Define a function to label representation as "High" or "Low"
label_rep_level <- function(pct) {
  if (is.na(pct)) {
    return(NA)
  } else if (pct > 0.5) {
    return("High")
  } else {
    return("Low")
  }
}

# Apply the function to create a new label
region_rep <- region_rep %>%
  mutate(women_rep_label = sapply(gender_women, label_rep_level))
```

```{r}
# Convert race columns into long format for later visualization.

race_long_format <- region_rep %>%
  pivot_longer(cols = starts_with("race_"),
               names_to = "race_group",
               values_to = "proportion")
```

```{r}
glimpse(race_long_format)
```

```{r}
# Calculate average attrition per gender by region.

region_attr %>%
  group_by(region) %>%
  summarise(
    avg_attr_women = mean(gender_women, na.rm = TRUE),
    avg_attr_men = mean(gender_men, na.rm = TRUE)
  )
```

```{r}
gender_rep_hiring %>%
  filter(!is.na(gender_women_rep), !is.na(gender_women_hire)) %>%
  group_by(region) %>%
  summarise(
    avg_rep = mean(gender_women_rep),
    avg_hire = mean(gender_women_hire)
  ) %>%
  ggplot(aes(x = avg_rep, y = avg_hire)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "gray40") +
  geom_text_repel(aes(label = region), size = 3) +
  labs(
    title = "Female Representation vs. Hiring by Region",
    x = "Average Female Representation",
    y = "Average Female Hiring Rate"
  ) +
  theme_minimal()
```

## Insights:
- A regression line is added to show overall correlation.
- EMEA is out performing The Americas and APAC. 
- We see Americas and APAC under the trend line.
- This shows female hiring rates in these areas are lagging behind.
- All three are below the soft 50% cap.

```{r}
race_long_format %>%
  filter(!is.na(proportion)) %>%
  group_by(region, workforce) %>%
  mutate(norm_prop = proportion / sum(proportion, na.rm = TRUE)) %>%
  ggplot(aes(x = region, y = norm_prop, fill = race_group)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ workforce) +
  labs(
    title = "Race Representation by Region and Workforce Type",
    x = "Region",
    y = "Proportion",
    fill = "Race Group"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Had to add the mutate to normalize to proper percentage...
```

## Insights:
- Each region shows distinct racial composition across workforce types.
- APAC heavily skewed to Asian representation.
- The Americas showing higher representation of White/European, Hispanic, and African.
- EMEA and leadership roles across regions show high White/European representation.
- MENA and Indigenous representation remains minimal as a whole.


```{r}
region_attr %>%
  pivot_longer(cols = starts_with("gender_"), names_to = "gender", values_to = "attrition_index") %>%
  ggplot(aes(x = report_year, y = attrition_index, color = gender, group = gender)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  facet_wrap(~ region) +
  labs(
    title = "Attrition Index Trends by Gender and Region",
    x = "Year",
    y = "Attrition Index",
    color = "Gender"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Insights:
- Across all regions, men consistently exhibit higher attrition index scores than woman.
- The gap between genders is widest in The Americas.
- APAC shows relatively stable attrition gaps overtime.
- EMEA's male attrition spiked pretty high in 2022, widening the gap.

```{r}
## Plotting the Global Self-ID. to find insights on these demographics.

ggplot(self_ID, aes(x = report_year, y = global, color = workforce, group = workforce)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
    title = "Global Self-ID Representation by Workforce Type",
    x = "Year",
    y = "Proportion (Self-Identified)",
    color = "Workforce"
  ) +
  theme_minimal()
```

## Insights:
- Self-identification rates have generally declined across all groups overall.
- LGBTQ+ employees made up the highest self-ID group in 2019 and kept this lead, even after decline.
- The drop could be a change in reporting practices, cultural stigma, or workforce demographic shifts.
- This trend shows the importance of psychological safety and transparent communication on Self-ID in the workplace.

```{r}
# Basic linear regression model
attrition_model <- lm(attrition_index ~ gender + region, data = region_attr %>%
                         pivot_longer(cols = starts_with("gender_"), names_to = "gender", values_to = "attrition_index"))

# View model summary
summary(attrition_model)
```

```{r}
# Expanding the basic model
# Prepare dataset
attrition_data <- region_attr %>%
  pivot_longer(cols = starts_with("gender_"), names_to = "gender", values_to = "attrition_index") %>%
  left_join(region_rep, by = c("region", "report_year", "workforce")) %>%
  mutate(workforce = as.factor(workforce),
         region = as.factor(region),
         gender = as.factor(gender))

# Fit model
attrition_model_v2 <- lm(attrition_index ~ gender + region + workforce + report_year, data = attrition_data)

# Summary
summary(attrition_model_v2)
```

## Machine Learning Model Insights:
- An expanded linear regression model was built to predict attrition index based on gender, region, workforce type, and report year. 
- The model explained approximately 64% of the variance in attrition index, suggesting a moderately strong fit.
- The analysis revealed that gender remains a highly significant predictor.
- Women had a 23 point lower attrition index compared to men, even after accounting for regional, workforce, and time differences. 
- No other factors (region, workforce, or year) showed statistically significant predictive power in this model.
- These results reinforce the critical need to consider gender dynamics in retention strategies. 
- However, additional variables such as tenure, compensation, and job level may further improve predictive power in future modeling efforts.

## Key Insights and Analysis

Through exploratory analysis and predictive modeling, several critical insights about global workforce diversity emerged:

- **Gender Representation and Hiring**: EMEA consistently outperforms the Americas and APAC regions in hiring women relative to their representation. However, no region reaches parity, highlighting persistent gender hiring gaps globally.
- **Racial Composition**: Across all workforce categories, White/European employees remain the most highly represented group, especially in leadership roles. Other racial and ethnic groups including Indigenous and MENA remain severely underrepresented.
- **Self-Identification Trends**: LGBTQ+ and disability self-identification rates have declined since 2019, suggesting a potential decrease in psychological safety or shifts in workforce demographics.
- **Attrition Trends by Gender**: Men exhibit higher attrition than women across all regions and workforce types. However, attrition patterns differ modestly by region over time.
- **Predictive Modeling**: A linear regression model predicting attrition index based on gender, region, workforce type, and report year explained approximately 64% of the variance. Gender emerged as the only statistically significant predictor. women showed a 23-point lower attrition index than men after controlling for other factors.

These findings highlight systemic gaps in workplace diversity and inclusion across regions and identity groups. They also suggest that gender specific retention strategies may be the most immediate opportunity for improving workforce equity.

## Recommendations

Based on the findings, several actions are recommended for organizational leadership:

- **Targeted Recruitment Initiatives**: Increase recruitment efforts for women in APAC and the Americas regions to close the gender hiring gap.
- **Leadership Development for Underrepresented Groups**: Invest in career advancement programs specifically for Black, Hispanic/Latinx, Indigenous, and MENA employees.
- **Enhance Psychological Safety**: Implement training and communication strategies to encourage safe, voluntary self-identification among employees.
- **Advanced Attrition Modeling**: Future workforce studies should include variables such as tenure, job level, salary, and performance scores to better predict retention risks.
