---
title: "Representation & Gender Diversity in Google’s Global Workforce"
output: html_notebook
---

## Guiding Research Question
- How does gender representation, hiring, and attrition vary across regions in Google's workforce?

## Data Access

**Primary Data**
- Region level workforce metrics such as representation, hiring, exits, and attrition.
- Source: Google Diversity Report from BigQuery & data available on this Github repo.
- - [Project Repository](https://github.com/PixelQuinn/Diversity-and-Inclusion-Analysis)

**Secondary Dataset:**
- Self-ID Representation data, including responses related to LGBTQ+ identity and disability status.
- Source: Internal Google workforce data (self-reported), accessed through Big Query.
- [Project Repository](https://github.com/PixelQuinn/Diversity-and-Inclusion-Analysis).

```{r}
library(tidyverse)
library(readr)
library(tidycensus)
library(janitor)
```

## Initial Inspection

```{r}
# Load data
region_rep <- read_csv("data-sets/region-non-intersectional-representation.csv") %>% clean_names()
region_hire <- read_csv("data-sets/region-non-intersectional-hiring.csv") %>% clean_names()
region_exit <- read_csv("data-sets/region-non-intersectional-exits-representation.csv") %>% clean_names()
region_attr <- read_csv("data-sets/region-non-intersectional-attrition-index.csv") %>% clean_names()

glimpse(region_rep)
summary(region_rep)
```

```{r}
self_ID <- read_csv("data-sets/selfid-representation.csv")

glimpse(self_ID)
```

```{r}
# Join the representation and hiring by region, year, and workforce type.

gender_rep_hiring <- region_rep %>%
  inner_join(region_hire, by = c("region", "report_year", "workforce"), suffix = c("_rep", "_hire"))
```

```{r}
gender_rep_hiring %>%
  filter(!is.na(gender_women_rep), !is.na(gender_women_hire)) %>%
  group_by(region) %>%
  summarise(
    avg_rep_women = mean(gender_women_rep),
    avg_hire_women = mean(gender_women_hire)
  )
```

```{r}
# Convert race columns into long format for later visualization.

race_long_format <- region_rep %>%
  pivot_longer(cols = starts_with("race_"),
               names_to = "race_group",
               values_to = "proportion")
```

```{r}
glimpse(race_long_format)
```
```{r}
colnames(region_attr_index)
```


```{r}
# Calculate average attrition per gender by region.

region_attr %>%
  group_by(region) %>%
  summarise(
    avg_attr_women = mean(gender_women, na.rm = TRUE),
    avg_attr_men = mean(gender_men, na.rm = TRUE)
  )
```

