---
title: "Representation & Gender Diversity in Google’s Global Workforce"
output: html_notebook
---

## Guiding Research Question
- How does gender representation, hiring, and attrition vary across regions in Google's workforce?

## Data Access

**Primary Data**
- Region level workforce metrics such as representation, hiring, exits, and attrition.
- Source: Google Diversity Report from BigQuery & data available on this Github repo.
- - [Project Repository](https://github.com/PixelQuinn/Diversity-and-Inclusion-Analysis)

**Secondary Dataset:**
- Self-ID Representation data, including responses related to LGBTQ+ identity and disability status.
- Source: Internal Google workforce data (self-reported), accessed through Big Query.
- [Project Repository](https://github.com/PixelQuinn/Diversity-and-Inclusion-Analysis).

```{r}
library(ggplot2)
library(tidyverse)
library(readr)
library(tidycensus)
library(janitor)
```

## Initial Inspection

```{r}
# Load data
region_rep <- read_csv("data-sets/region-non-intersectional-representation.csv") %>% clean_names()
region_hire <- read_csv("data-sets/region-non-intersectional-hiring.csv") %>% clean_names()
region_exit <- read_csv("data-sets/region-non-intersectional-exits-representation.csv") %>% clean_names()
region_attr <- read_csv("data-sets/region-non-intersectional-attrition-index.csv") %>% clean_names()

glimpse(region_rep)
summary(region_rep)
```

```{r}
self_ID <- read_csv("data-sets/selfid-representation.csv")

glimpse(self_ID)
```

```{r}
# Join the representation and hiring by region, year, and workforce type.

gender_rep_hiring <- region_rep %>%
  inner_join(region_hire, by = c("region", "report_year", "workforce"), suffix = c("_rep", "_hire"))
```

```{r}
gender_rep_hiring %>%
  filter(!is.na(gender_women_rep), !is.na(gender_women_hire)) %>%
  group_by(region) %>%
  summarise(
    avg_rep_women = mean(gender_women_rep),
    avg_hire_women = mean(gender_women_hire)
  )
```

```{r}
# Convert race columns into long format for later visualization.

race_long_format <- region_rep %>%
  pivot_longer(cols = starts_with("race_"),
               names_to = "race_group",
               values_to = "proportion")
```

```{r}
glimpse(race_long_format)
```

```{r}
# Calculate average attrition per gender by region.

region_attr %>%
  group_by(region) %>%
  summarise(
    avg_attr_women = mean(gender_women, na.rm = TRUE),
    avg_attr_men = mean(gender_men, na.rm = TRUE)
  )
```

```{r}
gender_rep_hiring %>%
  filter(!is.na(gender_women_rep), !is.na(gender_women_hire)) %>%
  group_by(region) %>%
  summarise(
    avg_rep = mean(gender_women_rep),
    avg_hire = mean(gender_women_hire)
  ) %>%
  ggplot(aes(x = avg_rep, y = avg_hire)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "gray40") +
  geom_text_repel(aes(label = region), size = 3) +
  labs(
    title = "Female Representation vs. Hiring by Region",
    x = "Average Female Representation",
    y = "Average Female Hiring Rate"
  ) +
  theme_minimal()
```

## Insights:
- A regression line is added to show overall correlation.
- EMEA is out performing The Americas and APAC. 
- We see Americas and APAC under the trend line.
- This shows female hiring rates in these areas are lagging behind.
- All three are below the soft 50% cap.

```{r}
race_long_format %>%
  filter(!is.na(proportion)) %>%
  group_by(region, workforce) %>%
  mutate(norm_prop = proportion / sum(proportion, na.rm = TRUE)) %>%
  ggplot(aes(x = region, y = norm_prop, fill = race_group)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ workforce) +
  labs(
    title = "Race Representation by Region and Workforce Type",
    x = "Region",
    y = "Proportion",
    fill = "Race Group"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Had to add the mutate to normalize to proper percentage...
```

## Insights:
- Each region shows distinct racial composition across workforce types.
- APAC heavily skewed to Asian representation.
- The Americas showing higher representation of White/European, Hispanic, and African.
- EMEA and leadership roles across regions show high White/European representation.
- MENA and Indigenous representation remains minimal as a whole.


```{r}
region_attr %>%
  pivot_longer(cols = starts_with("gender_"), names_to = "gender", values_to = "attrition_index") %>%
  ggplot(aes(x = report_year, y = attrition_index, color = gender, group = gender)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  facet_wrap(~ region) +
  labs(
    title = "Attrition Index Trends by Gender and Region",
    x = "Year",
    y = "Attrition Index",
    color = "Gender"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Insights:
- Across all regions, men consistently exhibit higher attrition index scores than woman.
- The gap between genders is widest in The Americas.
- APAC shows relatively stable attrition gaps overtime.
- EMEA's male attrition spiked pretty high in 2022, widening the gap.

```{r}
## Plotting the Global Self-ID. to find insights on these demographics.

ggplot(self_ID, aes(x = report_year, y = global, color = workforce, group = workforce)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
    title = "Global Self-ID Representation by Workforce Type",
    x = "Year",
    y = "Proportion (Self-Identified)",
    color = "Workforce"
  ) +
  theme_minimal()
```

## Insights:
- Self-identification rates have generally declined across all groups overall.
- LGBTQ+ employees made up the highest self-ID group in 2019 and kept this lead, even after decline.
- The drop could be a change in reporting practices, cultural stigma, or workforce demographic shifts.
- This trend shows the importance of psychological safety and transparent communication on Self-ID in the workplace.

## Closing thoughts:

This analysis set out to explore how gender, race, and self-identified identity groups are represented across regions and workforce types within a global organization. Through a mix of visual and statistical analysis, clear disparities were revealed across multiple dimensions of diversity.

Gender hiring and attrition trends varied quite a bit by region, with EMEA outperforming other regions in female hiring while APAC and The Americas showed constant gaps. Racial representation was highly uneven, with some groups, such as Indigenous and MENA employees, almost absent across all regions and roles. Self-ID trends showed an overall decline in disclosure, raising important questions about psychological safety and cultural transparency in the workplace.

These findings highlight the need for more regionally tailored diversity strategies, as well as stronger internal communication and support structures that empower employees to disclose identity safely and authentically.

As organizations continue to prioritize DEI, data-driven insights like these will be essential in not only measuring progress, but in guiding equitable, people centered actions.
