---
title: "Stat184 Class Poject - DEI Analysis"
output: html_notebook
---

**Research Question:**  
*How does gender and ethnicity representation, hiring, and attrition vary across different regions?*


### Primary Dataset
- **Source:** Google Diversity Annual Report data accessed through Google BigQuery
- **Collected by:** Google Inc.
- **When & Why:** Published annually to provide info around workforce demographics and support DEI efforts within the company
- **What is a case?** Each row represents a demographic group such as gender in a given region, along with a specific workforce metric such as hiring, exits, representation, or attrition
- **Variables of interest:**
- **Region: The global region of the workforce segment**
- **Gender or Ethnicity: The demographic group represented**
- **Hiring, Representation, Exits, Attrition Index: The workforce metric value, depending on the dataset**



### Secondary Dataset

- **Source:** American Community Survey (ACS) 5-Year Estimates accessed via the `tidycensus` package in R
- **Collected by:** U.S. Census Bureau
- **When & Why:** The ACS is collected annually to provide critical demographic, social, economic, and housing data across U.S. regions. This data helps contextualize diversity metrics, such as gender and racial/ethnic representation, in broader U.S. labor markets.
- **What is a case?** Each case represents a region (e.g., West, Midwest, South, Northeast) or state, along with aggregated population estimates by race, gender, and education level.
- **Number of Cases:** Varies depending on granularity (e.g., ~50 states or 4 census regions), filtered down based on focus
- **Variables of interest:**  
  - Total population by gender  
  - Total population by race/ethnicity  
  - Percent of population with a bachelor’s degree or higher (for pipeline comparison)

```{r}
library(tidyverse)
library(ggplot2)
library(tidycensus)
```

```{r}
#API for Tidycensus,PERSONAL REMINDER TO HIDE FROM PROD!!!!!
census_api_key("MY API HERE", install = TRUE) #Hiding upon submitting for safety & best practice reasons.
```


```{r}
# Load region-level workforce metrics
region_rep <- read_csv("data-sets/region-non-intersectional-representation.csv")
region_hire <- read_csv("data-sets/region-non-intersectional-hiring.csv")
region_exit <- read_csv("data-sets/region-non-intersectional-exits-representation.csv")
region_attrition <- read_csv("data-sets/region-non-intersectional-attrition-index.csv")
```

```{r}
glimpse(region_rep)
summary(region_rep)
head(region_rep)
```
```{r}
glimpse(region_hire)
summary(region_hire)
head(region_hire)
```
```{r}
glimpse(region_exit)
summary(region_exit)
head(region_exit)
```
```{r}
glimpse(region_attrition)
summary(region_attrition)
head(region_attrition)
```
```{r}
region_rep %>%
  ggplot(aes(x = region, y = gender_women, fill = report_year)) +
  geom_col(position = "dodge") +
  facet_wrap(~ workforce) +
  labs( title = "Representation of Women by Region and Workforce", x = "Region", y = "Proportion of Women") +
  theme_minimal()
```
```{r}
region_hire <- region_hire %>%
  mutate(report_year = as.factor(report_year))

ggplot(region_hire, aes(x = region, y = gender_women, fill = report_year)) +
  geom_col(position = "dodge") +
  facet_wrap(~ workforce) +
  labs(title = "Hiring Rate of Women by Region and Workforce Segment",
       x = "Region", y = "Proportion of Women Hired") +
  theme_minimal()
```
```{r}
region_exit %>%
  ggplot(aes(x = region, y = gender_women, fill = report_year)) +
  geom_col(position = "dodge") +
  facet_wrap(~ workforce) +
  labs(
    title = "Exit Rate of Women by Region and Workforce Segment",
    x = "Region", y = "Proportion of Women Exiting"
  ) +
  theme_minimal()
```

### Summary of Preliminary Observations and Intuition

This exploratory data analysis investigates how gender representation, hiring, and attrition differ across global regions within Google’s workforce. The dataset, sourced from Google’s Diversity Annual Report through BigQuery, includes metrics for tech, non-tech, leadership, and overall segments, broken down by region. Initial EDA will be based off of a soft 50% benchmark as well as comparing the data to eachother.

#### Key Observations:
- **Representation:**
  - Women remain underrepresented in tech and leadership roles across all regions.
  - Non-tech roles show near gender parity, particularly in APAC and EMEA.
  - Minimal progress had been made between 2021 and 2022 overall.

- **Hiring:**
  - Hiring of women in leadership roles improved significantly in EMEA and APAC from 2021 to 2022.
  - Non-tech hiring rates were the highest, typically surpassing the 45% mark.
  - Tech hiring rates are the lowest and relatively stagnant. This shows potential pipeline issues in this field.
 
- **Exits:**
  - Women in non-tech roles exit at the highest rates, usually around 40% or greater.
  - Tech exit rates were lowest, but increase steadily and noticeably in 2022.
  - Exit patterns show potential retention challenges, especially in non-tech roles.
 
- **Post EDA Thoughts:**
  Diversity is beyond just hiring the target variable. It is about hiring the right candidate and also retaining them. We need to identify patterns such as in **Non-tech exits** to find issues and insights on how to reconcile. 
As well as **tech hiring** potentially showing a pattern for a possible internal challenge as well. While progress is being made over all, the slow progress in tech is combatted by a steady exit for women showing some barriers within the role,
hurting long term impact. This lays the foundation for a deeper analysis exploring the data further.
