---
title: "Stat184 Class Poject - DEI Analysis"
output: html_notebook
---

**Research Question:**  
*How does gender and ethnicity representation, hiring, and attrition vary across different regions?*


### Primary Dataset
- **Source:** Google Diversity Annual Report data accessed through Google BigQuery
- **Collected by:** Google Inc.
- **When & Why:** Published annually to provide info around workforce demographics and support DEI efforts within the company
- **What is a case?** Each row represents a demographic group such as gender in a given region, along with a specific workforce metric such as hiring, exits, representation, or attrition
- **Variables of interest:**
- **Region: The global region of the workforce segment**
- **Gender or Ethnicity: The demographic group represented**
- **Hiring, Representation, Exits, Attrition Index: The workforce metric value, depending on the dataset**



### Secondary Dataset

- **Source:** American Community Survey (ACS) 5-Year Estimates accessed via the `tidycensus` package in R
- **Collected by:** U.S. Census Bureau
- **When & Why:** The ACS is collected annually to provide critical demographic, social, economic, and housing data across U.S. regions. This data helps contextualize diversity metrics, such as gender and racial/ethnic representation, in broader U.S. labor markets.
- **What is a case?** Each case represents a region (e.g., West, Midwest, South, Northeast) or state, along with aggregated population estimates by race, gender, and education level.
- **Number of Cases:** Varies depending on granularity (e.g., ~50 states or 4 census regions), filtered down based on focus
- **Variables of interest:**  
  - Total population by gender  
  - Total population by race/ethnicity  
  - Percent of population with a bachelor’s degree or higher (for pipeline comparison)

```{r}
library(tidyverse)
library(ggplot2)
library(tidycensus)
```

```{r}
#API for Tidycensus,PERSONAL REMINDER TO HIDE FROM PROD!!!!!
census_api_key("MY API HERE", install = TRUE) #Hiding upon submitting for safety & best practice reasons.
```


```{r}
# Load region-level workforce metrics
region_rep <- read_csv("data-sets/region-non-intersectional-representation.csv")
region_hire <- read_csv("data-sets/region-non-intersectional-hiring.csv")
region_exit <- read_csv("data-sets/region-non-intersectional-exits-representation.csv")
region_attrition <- read_csv("data-sets/region-non-intersectional-attrition-index.csv")
```

```{r}
glimpse(region_rep)
summary(region_rep)
head(region_rep)
```
```{r}
glimpse(region_hire)
summary(region_hire)
head(region_hire)
```
```{r}
glimpse(region_exit)
summary(region_exit)
head(region_exit)
```
```{r}
glimpse(region_attrition)
summary(region_attrition)
head(region_attrition)
```
```{r}
region_rep %>%
  ggplot(aes(x = region, y = gender_women, fill = report_year)) +
  geom_col(position = "dodge") +
  facet_wrap(~ workforce) +
  labs( title = "Representation of Women by Region and Workforce", x = "Region", y = "Proportion of Women") +
  theme_minimal()
```
```{r}
region_hire <- region_hire %>%
  mutate(report_year = as.factor(report_year))

ggplot(region_hire, aes(x = region, y = gender_women, fill = report_year)) +
  geom_col(position = "dodge") +
  facet_wrap(~ workforce) +
  labs(title = "Hiring Rate of Women by Region and Workforce Segment",
       x = "Region", y = "Proportion of Women Hired") +
  theme_minimal()
```
```{r}
region_exit %>%
  ggplot(aes(x = region, y = gender_women, fill = report_year)) +
  geom_col(position = "dodge") +
  facet_wrap(~ workforce) +
  labs(
    title = "Exit Rate of Women by Region and Workforce Segment",
    x = "Region", y = "Proportion of Women Exiting"
  ) +
  theme_minimal()
```

